# Results

```{r include=FALSE}
library(tidyverse)
library(ggplot2)
library(GGally)
library(vcd)
library(RColorBrewer)
library(ggalluvial)
#devtools::install_github("timelyportfolio/parcoords")
library(parcoords)
library(pander)
```

```{r}
shot_data = read_csv("shotchartdetail.csv")
```


In this chapter, we will our process to come up with the answers and our conclusions to each of the questions we proposed in the introduction section. Here are our questions proposed in the introduction chapter:

1. (On team level)
What is the distribution of the total number of shots made for all players in the entire season? What about shot probability? Is there any difference between shots attempted and goaled for different teams in different zones? What about differences between the home team and visiting team?

2. (On individual player level)
Whether Russ Westbrook (Guard) has different shooting behavior (the shooting count and shot probability on different areas on the field) in different periods of the game? What about LeBron James (Forward)? Is LeBron’s shooting behavior similar to Russ ’s in different periods of the game?

Note that in the following analysis, we need to plot the NBA court to visualize the shooting data. The code for drawing the court is adapted from https://github.com/toddwschneider/ballr, and the court graph is as follows:

```{r}
circle_points = function(center = c(0, 0), radius = 1, npoints = 360) {
  angles = seq(0, 2 * pi, length.out = npoints)
  return(data_frame(x = center[1] + radius * cos(angles),
                    y = center[2] + radius * sin(angles)))
}

width = 50
height = 94 / 2
key_height = 19
inner_key_width = 12
outer_key_width = 16
backboard_width = 6
backboard_offset = 4
neck_length = 0.5
hoop_radius = 0.75
hoop_center_y = backboard_offset + neck_length + hoop_radius
three_point_radius = 23.75
three_point_side_radius = 22
three_point_side_height = 14
court_themes = list(
  light = list(
    court = 'floralwhite',
    lines = '#999999',
    text = '#222222',
    made = '#00bfc4',
    missed = '#f8766d',
    hex_border_size = 1,
    hex_border_color = "#000000"
  ),
  dark = list(
    court = '#000004',
    lines = '#999999',
    text = '#f0f0f0',
    made = '#00bfc4',
    missed = '#f8766d',
    hex_border_size = 0,
    hex_border_color = "#000000"
  )
)


plot_court = function(court_theme = court_themes$light, use_short_three = FALSE) {
  if (use_short_three) {
    three_point_radius = 22
    three_point_side_height = 0
  }
  
  court_points = tibble(
    x = c(width / 2, width / 2, -width / 2, -width / 2, width / 2),
    y = c(height, 0, 0, height, height),
    desc = "perimeter"
  )
  
  court_points = bind_rows(court_points , tibble(
    x = c(outer_key_width / 2, outer_key_width / 2, -outer_key_width / 2, -outer_key_width / 2),
    y = c(0, key_height, key_height, 0),
    desc = "outer_key"
  ))
  
  court_points = bind_rows(court_points , tibble(
    x = c(-backboard_width / 2, backboard_width / 2),
    y = c(backboard_offset, backboard_offset),
    desc = "backboard"
  ))
  
  court_points = bind_rows(court_points , tibble(
    x = c(0, 0), y = c(backboard_offset, backboard_offset + neck_length), desc = "neck"
  ))
  
  foul_circle = circle_points(center = c(0, key_height), radius = inner_key_width / 2)
  
  foul_circle_top = filter(foul_circle, y > key_height) %>%
    mutate(desc = "foul_circle_top")
  
  foul_circle_bottom = filter(foul_circle, y < key_height) %>%
    mutate(
      angle = atan((y - key_height) / x) * 180 / pi,
      angle_group = floor((angle - 5.625) / 11.25),
      desc = paste0("foul_circle_bottom_", angle_group)
    ) %>%
    filter(angle_group %% 2 == 0) %>%
    select(x, y, desc)
  
  hoop = circle_points(center = c(0, hoop_center_y), radius = hoop_radius) %>%
    mutate(desc = "hoop")
  
  restricted = circle_points(center = c(0, hoop_center_y), radius = 4) %>%
    filter(y >= hoop_center_y) %>%
    mutate(desc = "restricted")
  
  three_point_circle = circle_points(center = c(0, hoop_center_y), radius = three_point_radius) %>%
    filter(y >= three_point_side_height, y >= hoop_center_y)
  
  three_point_line = tibble(
    x = c(three_point_side_radius, three_point_side_radius, three_point_circle$x, -three_point_side_radius, -three_point_side_radius),
    y = c(0, three_point_side_height, three_point_circle$y, three_point_side_height, 0),
    desc = "three_point_line"
  )
  
  court_points = bind_rows(
    court_points,
    foul_circle_top,
    foul_circle_bottom,
    hoop,
    restricted,
    three_point_line
  )
  
  
  court_points <- court_points
  
  ggplot() +
    geom_path(
      data = court_points,
      aes(x = x, y = y, group = desc),
      color = court_theme$lines
    ) +
    coord_fixed(ylim = c(0, 45), xlim = c(-25, 25)) +
    theme_minimal(base_size = 22) +
    theme(
      text = element_text(color = court_theme$text),
      plot.background = element_rect(fill = 'floralwhite', color = 'floralwhite'),
      panel.background = element_rect(fill = court_theme$court, color = court_theme$court),
      panel.grid = element_blank(),
      panel.border = element_blank(),
      axis.text = element_blank(),
      axis.title = element_blank(),
      axis.ticks = element_blank(),
      legend.background = element_rect(fill = court_theme$court, color = court_theme$court),
      legend.margin = margin(-1, 0, 0, 0, unit = "lines"),
      legend.position = "bottom",
      legend.key = element_blank(),
      legend.text = element_text(size = rel(1.0))
    )
}

plot_court(court_themes$light)
```

## Team Shooting Pattern Analysis

Firstly, to get a general understanding of how many 2-point and 3-point shots each team attempted for season 2020-21, we create a Cleveland Dot Plot as follows:

```{r}
shot_data %>%
  select(team_name, shot_type) %>%
  group_by(team_name,shot_type) %>%
  summarize(n = n(),.groups = 'drop') %>%
  pivot_wider(values_from = n, names_from = shot_type) %>%
  rename_all(make.names) %>%
  rename(two_point_shots = X2PT.Field.Goal, three_point_shots = X3PT.Field.Goal) %>%
  mutate(shots_sum = two_point_shots+ three_point_shots) %>% # calculate total
  arrange(desc(shots_sum)) %>% # order by total
  pivot_longer(cols = c("two_point_shots", "three_point_shots","shots_sum")) %>% # tidy the data
  ggplot(aes(x = value, y = reorder(team_name, value), color = name))+
  geom_point()+
  theme_linedraw()+ 
  labs(title = "Cleveland Dot Plot for Attempted Shots Count for Each Team", x = "count", y = "Team Name")
```

From the Cleveland Dot Plot, we can see that the total shots made by each team for the entire season doesn't vary a lot, ranging from around 6000 to 6500. However, for different teams, they have totally different preference for 2-point shots and 3-point shots, which may be resulted by types of players in that team. 

For example, Utah Jazz has almost same counts for 2-point shots and 3-point shots with the most 3-point shots in the league. On the contrary, San Antonio Spurs has least count of 3-point shots and most count of 2-point shots in the league. We could guess that San Antonio Spurs doesn't have a lot of 3-point players in the team so they prefer goaling by 2-point shots.

After getting a general understanding of types and shots attempted for each team, now we want to focus on the Field Goal Percentage (total shots goaled/ total shots attempted). 

```{r}
# order teams based on total shot probability
order <- shot_data %>%
  select(team_name,shot_attempted_flag,shot_made_flag) %>%
  group_by(team_name) %>%
  summarize(attempts = sum(shot_attempted_flag),
            goals = sum(shot_made_flag)) %>%
  mutate(goal_prob = goals/attempts)  %>%
  arrange(desc(goal_prob)) %>%
  pull(team_name)

shot_data %>%
  select(team_name,shot_zone_range,shot_attempted_flag,shot_made_flag) %>%
  mutate(team_name = factor(team_name, levels = order)) %>%
  mutate(shot_zone_range = factor(shot_zone_range, levels = c("Less Than 8 ft.", "8-16 ft.", "16-24 ft.","24+ ft.","Back Court Shot"))) %>%
  group_by(team_name,shot_zone_range) %>%
  summarize(attempts = sum(shot_attempted_flag),
            goals = sum(shot_made_flag)) %>%
  mutate(goal_prob = goals/attempts)  %>%
  ggplot(aes(x = team_name, y = goal_prob, color = shot_zone_range, shape = shot_zone_range))+
  geom_point()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  labs(title = "Scatter Plot of Goal Probability for Each Team", x = "Team", y = "Goal%")
```

## Individual Player Analysis
Here we will specially explore 

